---
title: ClustOmics
author: Evan Stene, Shahab Helmi, Lucas Gillenwater, Andrew Hill, Yonghua Zhuang
date: "`r format(Sys.Date())`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide", message = FALSE, eval = FALSE, warning = FALSE)
library(openxlsx)
library(roxygen2)
library(devtools)
install("./ClustOmics")
library(ClustOmics)

```


# ClustOmics
ClustOmics is an R package for implementing the entire pipeline reported in (citation)

#### NB - The following sections are required for submission to bioconductor

### Installation
### Documentation
* tutorial
* reference manual
* News

### Details
* biocViews 
	-"biocViews terms are “keywords” used to describe a given package. They are broadly divided into three categories, representing the type of packages present in the Bioconductor Project - Software, Annotation Data, Experiment Data"
* License 
* Dependencies
* Vignette
* Imports
* LinkingTo
* Suggests 
* SystemRequirements
* Enhances
* URL
* BugReports
* Depends On Me
* Imports Me
* Suggests me
* Links to Me

### Package Archives
* Source Package
* Windows Binary
* Mac OS X
* Source Repository
* Source Repository (Developer Access)
* Package Short Url
* Package Downloads Report
* Old Source Packages for BioC 3.10

# Data Input
* Loading in necessary data
* 12/5/2019 - Only loading in the proteomic data and important clinical variables


```{r Data}
###### This could be cleaned to be streamlined within the filter function
##### Datasets
clinical_file <- "/Bowler/home/gillenwaterl/COPDGene/Data/clinical/COPDGene_P1P2_All_Visit_29sep18.txt"
clinical_variables_file <- "/Bowler/home/gillenwaterl/COPDGene/Data/clinical/P1P2_Pheno_w_QCT_DataDict_annotated_29sep18_important clinical variables_norace_gender_age.xlsx"
protein_file <- "/Bowler/home/gillenwaterl/COPDGene/Data/somalogic/COPDGene_SOMA1300_P1P2_Sep18.txt"

# All datasets should have SIDs as rownames and features 
# All clinical data
clinic <- read.delim(clinical_file)
# # Important clinical variables
imp <- read.xlsx(clinical_variables_file)
# creating separate datasets for each phase without never smokers or those with exclusionary diseases
clinic2 <- clinic[clinic$visitnum == 2 & clinic$smoking_status > 0 & clinic$ExclusionaryDisease == "",]
rownames(clinic2) <- clinic2$sid
clinic2 <- clinic2[,  names(clinic2) %in% imp$VariableName]



### Somalogic
prot <- read.delim(protein_file)
prot <- prot[prot$Visitnum == 2,]
rownames(prot) <- prot$SID
prot <- prot[, 4:ncol(prot)]



```

# Filtering and adjustment
* Purpose: 
	- Data is adjusted for potential confounding factors (blood counts)
	- The filtering function is a utility for reducing the feature set entering the dimension reduction function and removing "noisy" molecules. 
	
* Input: 
	- Omic data that have been previously normalized and adjusted for potential batch effects. 
	- Clinical data subsetted to only include interesting variables chosen by a clinician (Dr. Russell Bowler)

* Processes performed: 
	- Log transform the data
	- Adjust for confounding variables (blood counts)
	- Center and Scale (standardize)
	- Filter based on associations with important clinical variabless

* Output: 
	- Filtered Dataset with SIDs as rows and features as columns
	- List of clinical features significantly (beyond set threshold) associated with each feature
	- Other visualizations? 


```{r filtering and adjustment}
# Log transform the data 
prot <- log(prot)

# Adjusting the data for potential confounders
# vector of variables to adjust for
vars <- c("hemoglobin", "neutrophl_pct", "lymphcyt_pct", "eosinphl_pct", "monocyt_pct")
adjusted.data <- linear_adjust(prot, clinic2, vars)

# Standardize data
adjusted.data$standardized <- center_and_scale(adjusted.data$residuals)

# further remove the blood count variables (CBC) and BMI from filtering clincial variables
adjusted.data$clinical_noCBC_noBMI <- adjusted.data$clinical[, !(names(adjusted.data$clinical) %in% c("wbc","lymphcyt","lymphcyt_pct", "neutrophl_pct", "neutrophl", "monocyt_pct",  "monocyt", "eosinphl_pct", "eosinphl", "hemoglobin", "hematocrit", "MCV", "MCH", "MCHC", "Platelets", "RBC", "basophl", "basophl_pct", "BMI")) ]
# Filtered data- would be good to add output of features and their associations here (LG 12/6/19)
filtered.data <- filter_function(dat = adjusted.data$standardized, 
								clinical = adjusted.data$clinical_noCBC_noBMI, 
								padjust = "fdr", 
								sig_val = 5e-5,
								filter_function = "non-parametric", 
								parallel = TRUE,



``` 


# Dimension Reduction
* Notes 
	- need library keras
	- need to install tensor flow. How to ensure that those using the package have it?
	- 

```{r autoencoder}

install("./ClustOmics")
library(ClustOmics)

auto.data <- autoencoder(df = filtered.data$filtered, 
						 layer_sizes = c(128, 64, 16), 
                        pretrain_epochs = 100, finetune_epochs = 200,
                        pretrain_lr = 0.01, finetune_lr = 0.00001,
                        batch_size = 100, validation_split = 0.2,
                        use_gpu = FALSE)

```


# Clustering
# Clinical Association
# Feature Selection














